Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints

### 2017-09-30 2017/ParOptBook/Chapter4.tex
parOptBook: finish main work on chapter 4

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                           % not available on your system

\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

%custom package
\usepackage{mathtools}

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

@@

\example{
\label{ex4:problem}
Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6;2.2],m=2$,
\begin{gather*}
  \varphi(x)=\cos(18x-3)\sin(10x-7)+1.5, \\
  g_1(x)=\exp(-x/2)\sin(6x-1.5), \\
  g_2(x)=|x|\sin(2\pi x-0.5).
\end{gather*}
he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
\begin{figure}[h]
  \label{fig:4_1}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_1.png}
  \caption{One-dimensional global optimization problem}
\end{figure}
}

@@

\begin{figure}[h]
  \label{fig:4_2}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_2.png}
  \caption{A problem with partially computable functions}
\end{figure}

@@

\begin{figure}[h]
  \label{fig:4_3}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_3.png}
  \caption{An ''index'' function}
\end{figure}

Since in the general case the problem \eqref{eq4:problem2} may have no solution (i.e., the feasible subdomain $Q_{m+1}$ might appear to be empty because of the incompatibility of the constrains) an auxiliary problem always having a solution is associated with it. Since conditions \eqref{eq4:condition} are equivalent to the conditions
\begin{displaymath}
  x\in Q_\nu,x\not\in Q_{\nu+1},
\end{displaymath}
this auxiliary problem can be written in the form
\begin{equation}
  \label{eq4:problem3}
  g_M^*=g_M(x_M^*)=\min\{g_M(x):x\in Q_M\},
\end{equation}
where $M$ is the greatest possible value of the index i. e.
\begin{equation}
  1\le M=\max\{\nu(x):x\in[a,b]\}\le m+1
\end{equation}
Since the set $Q_M$ is always nonempty, problem \eqref{eq4:problem3} always has a solution. When $M=m+1$ the solution $x^*=x^*_m+1$  is also the solution of the initial problem \eqref{eq4:problem}. When $M<m+1$ the inequality $g_M^*<0$ fulfilled necessarily can be used as an indicator of the incompatibility of the constraints.

@@

As a result, the arcs of $\psi(x)$ will be Lipschitzian with the constant $L=1$ in each subdomain $Q_\nu,1\le\nu\le M$. A graph of the function $\psi(x)$ for the example (\ref{ex4:problem}) is presented in Fig. \ref{fig:4_4}. This new function will have discontinuities of the first type at the boundary points of the subdomains $Q_\nu$ from \eqref{eq4:q}. Nevertheless, one can estimate the global minimizer $x^*_M$ using the results of $k$ trials for \eqref{eq4:trial} at the points $x_1,\dots, x_k$ in $[a,b]$ (see \cite{3}).

@@

For example, a case for $k=4$ is presented in Fig. \ref{fig:4_4}. The union of the segments highlighted with a bold line in Fig. \ref{fig:4_4} is a complement of the set from the right hand side of \eqref{eq4:lip_opt_est} and does not include the optimum point.
\begin{figure}[h]
  \label{fig:4_4}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_4.png}
  \caption{An estimate of the optimum point}
\end{figure}

@@

This theorem is given without a proof, which can be found, for example, in \cite{}[1, 2]. Various modifications of this algorithm and the corresponding theory of convergence are given in \cite{}[5–10].

@@

\begin{figure}[h]
  \label{fig:4_5}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_5.png}
  \caption{Minimization by means of Penalty Function Method}
\end{figure}

@@

Here $k_i$ is the number of computations of the $i$ problem function, i.e., of the function $g_i(x )$.
\begin{figure}[h]
  \label{fig:4_6}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_6.png}
  \caption{The minimization of an ''index'' function}
\end{figure}

\section{Index algorithm taking into account the existence of $\epsilon$-reserved solutions}
Let us return to the consideration of a one-dimensional problem of the kind \eqref{eq4:problem}
\begin{displaymath}
  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,1\le j\le m\}
\end{displaymath}
In the case when the objective function and the left-side parts of the constraints are partially computable, it is necessary to use the form \eqref{eq4:problem2}.
\example
{
Let us consider problem \eqref{eq4:problem} with $x\in[0,6,2,2]$, $m=3$ as an illustration:
\begin{gather*}
\varphi(x)=\cos(18x-3)\sin(10x-7)+1, \\
g_1(x)=\exp(-x/2)\sin(6x-1.5) \\
g_2(x)=\sin(4x-2.2)+\cos(6x-2.9) \\
g_3(x)=|x|\sin(2\pi x - 0.5).
\end{gather*}
}
\begin{figure}[h]
  \label{fig:4_7}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_7.jpg}
  \caption{The solution of a problem by Index Method}
\end{figure}
The functions $g_j , 1\le j\le 4$, are pictured in Fig. \ref{fig:4_7}. Assuming a partial calculability, the arcs of
the functions corresponding to the domains $Q_j ,1\le j\le 4$, from \eqref{eq4:q} are presented in Fig. \ref{fig:4_8}.

\begin{figure}[h]
  \label{fig:4_8}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_8.jpg}
  \caption{The problem solving by Index Method with the $\epsilon$-reservs}
\end{figure}

@@

As it has been demonstrated by the experiment with solving the example described above, the increase of the reserve values may result in an essential decrease of the trial concentration in the boundary points of non-feasible domains. The theoretical properties of the trial sequences generated by the index algorithm are determined by the following statement (see the proof in \cite{}[4] also).
\begin{theorem}
  Assume the convergence conditions for the index algorithm to be satisfied and the inequalities
  \begin{displaymath}
    g_\nu(x)\ge\delta_\nu,x\in[\alpha,\beta],\delta_\nu\ge 0
  \end{displaymath}
  to be satisfied in a subinterval $[\alpha,\beta]\subset Q_\nu \: 1\le\nu\le m$, the length of which meets the condition
  \begin{displaymath}
    \beta-\alpha>\frac{2(r_\nu-1)(\delta_\nu+\epsilon_\nu)}{r^2_\nu L_\nu}.
  \end{displaymath}

@@

\begin{figure}[h]
  \label{fig:4_9}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_9.png}
  \caption{The operation characteristics in class $G_{SH}$}
\end{figure}

\begin{figure}[h]
  \label{fig:4_10}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_10.png}
  \caption{The operation characteristics in class $G_{HL}$}
\end{figure}

The operational characteristics for both method obtained for the test problem sets generated by $G_{SH}$ and $G_{HL}$ are presented in Figures 4.9 and 4.10, respectively. The lower (solid) curves refer to the initial index algorithm while the upper (dashed) ones correspond to the index algorithm with the adaptive reserves. Such arrangement of the curves demonstrates the algorithm with the
adaptive reserves to provide much faster obtaining the estimates falling into given vicinity of the solution than the initial index algorithm. The average numbers of iterations for the method
without the reserves and for the method with the adaptive reserves were equal to 171 and 117 for the problems $G_{SH}$ and 199 and 142 for the ones $G_{HL}$ , respectively.

@@

\textbf{Proof} of the theorem is given in \cite{}[5].

@@

\begin{figure}[h]
  \label{fig:4_11}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_11.jpg}
  \caption{The distribution of problems of the class $G_{SH}$}
\end{figure}

\begin{figure}[h]
  \label{fig:4_12}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_12.jpg}
  \caption{The distribution of problems of the class $G_{HL}$}
\end{figure}

The distribution of the interval lengths $[\alpha,\beta]$ from (4.41) including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig. 4.11. Fig. 4.12 demonstrates a similar distribution for the problems of the second set. In these figures, the values $(\beta=\alpha)/(b-a)$ (in percentage) are plotted against the abscissa axes while the numbers of problems in the corresponding sets are plotted against the ordinate ones. In order to obtain the discussed estimates, the lengths of the intervals $[\alpha,\beta]$ corresponding to the respective functions were determined by a numerical method.

@@

The constraint functions $g_\nu( x ),\: 1 \le\nu\le 5$, and the objective function $\varphi (x)$ for a problem from the class $G_{HP}$ are plotted in Fig. 4.13 as an example. The feasible domains which are formed by the separate constraints are marked on the abscissa axis. The whole feasible domain of the problem (the intersection of all feasible domains of the constraints) is shown along with the graph of the objective function.

\begin{figure}[h]
  \label{fig:4_13}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_13.png}
  \caption{Function plots of a test problem from the class $G_{HL}$}
\end{figure}

@@

\begin{figure}[h]
  \label{fig:4_14}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_13.png}
  \caption{Operational characteristics for the class $G_{SH}$}
\end{figure}

\begin{figure}[h]
  \label{fig:4_15}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_15.png}
  \caption{Operational characteristics for the class $G_{HL}$}
\end{figure}

The operational characteristics for both the methods obtained on the sets produced by generators $G_{SH}$ and $G_{HL}$ are presented in figures 4.14 and 4.15, respectively. The lower (solid) curves correspond to FCOM while the upper (dashed) ones --- to ACOM. These positions of the curves demonstrate that the algorithm with the adaptive check order provides in average faster obtaining the estimates falling into given vicinity of the solution than the algorithm with the fixed checking order.

\begin{figure}[h]
  \label{fig:4_16}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_16.png}
  \caption{Average number of function values computations for the class $G_{SH}$}
\end{figure}

\begin{figure}[h]
  \label{fig:4_17}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_17.png}
  \caption{Average number of function values computations for the class $G_{HL}$}
\end{figure}

@@

  It follows from the trial execution rules and condition (4.35) that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive. Therefore, according to \eqref{eq4:z_const_par},
  \begin{displaymath}
    z_i=g_\nu(x_i)\ge z^*_nu,\:1\le\nu=\nu(x_i)\le m+1,\: 0\le i\le k.
  \end{displaymath}
    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit (4.56), taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality (4.57), one can derive out the truth of statement (4.58).
\end{proof}
\begin{lemma}
  (on the existence of a feasible point). Assume the conditions of Theorem \ref{th4:convergence_par} to be fulfilled. Then, there exists the trial index $h\ge 1$, for which
  \begin{equation}
    \label{eq4:62}
    \nu(x^h)=m+1.
  \end{equation}
\end{lemma}
\begin{proof}
  Let us assume the opposite and consider the interval $[x_{t-1},x_t]$ including the problem solution $x^*$ from \eqref{eq4:problem} at the step $k$, i.e., $t=t(k)$. Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from (4.52) should satisfy the conditions
  \begin{equation}
    x^*\in [\alpha,\beta]\subset[x_{t-1}, x_t].
  \end{equation}
  Then, from this condition as well as from the Lipschitz properties of the problem functions the relations
  \begin{gather}
    z_{t-1}\le L_\nu(\alpha - x_{t-1})\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t-1}) \\
    z_t\le L_\nu(x_{t}-\beta)\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t})
  \end{gather}
  follow. It was taken into account when deriving out of the above relations that
  \begin{displaymath}
    g_\nu(x)\le 0,\: x\in[\alpha,\beta]\subset Q.
  \end{displaymath}
  Analogously, the case $\nu=\nu(x_{t-1})=\nu(x_t)$ corresponds to the inequality
  \begin{equation}
    z_t+z_{t-1}\le L_\nu \Delta_t - L_\nu(\beta-\alpha).
  \end{equation}

  Now, from the characteristic rules \eqref{eq4:characteristic_par}, from inequalities (4.64)–(4.66), and from the inequality
  \begin{displaymath}
    z^*_\nu\ge-\epsilon_\nu
  \end{displaymath}
  resulting from \eqref{eq4:z_const_par} either the estimate
  \begin{displaymath}
    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+2(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
  \end{displaymath}
  or the estimate
  \begin{displaymath}
    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+4(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
  \end{displaymath}
  is true. Taking into account the third condition of Theorem \ref{th4:convergence_par} and the fourth one, in both cases we get the inequality
  \begin{equation}
    R(t(k))>0
  \end{equation}
  which holds for sufficiently large step numbers $k$.

  From (4.58), (4.67), and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from (4.63) follows. Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals. After that, at least one of those will include the point $x^*$. The repetition of this division generates (with $k\to\infty$) an infinite sequence of the subintervals shrinking to the point $x^*$ . In other words, if the third condition of Theorem \ref{th4:convergence_par} and the fourth one are satisfied, $x^*$ is the limit point of the sequence ${x^k}$.

@@

\begin{lemma}
  on the feasibility of any limit point). Assume the conditions of Theorem \ref{th4:convergence_par} to be met. Then any limit point $\overline x$ of the sequence $\{x_k\}$ is feasible, i.e.,
  \begin{equation}
    \label{eq4:68}
    \nu(\overline x)=m+1
  \end{equation}
\end{lemma}

@@

  According to the trial execution rules and to conditions (4.35) and (4.72), for any interval from (4.70) the relations
  \begin{displaymath}
    \omega_1=\nu(x_{t-1})\in M,\: \omega_2=\nu(x_t)\in N,\:t=t(q^s),
  \end{displaymath}
  hold where $1\le \omega_1,\: \omega_2\le m$. Moreover, according to the rules (4.39) and (4.40) defining the constraints’ check order at trial points,
  \begin{displaymath}
    \nu(x_{t-1}),\:\nu(x_t)\in\{\omega_1,\omega2\},\: t=t(q^{p+1}).
  \end{displaymath}
  In other words, for sufficiently large number $s>1$ the indices assigned to the boundary points of the considered system of intervals $[x_{t-1},x_t],\: t=t(q^s ): s\ge S$ can take the values from the set $\{\omega_1,\omega2\}$ only.

@@

\end{proof}

>>>>

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
%                           % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%%custom package
%\usepackage{mathtools}
%\usepackage{cite}
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{document}

@@

\example{
\label{ex4:problem}
Let us consider a problem of type \eqref{eq4:problem} where $x\in [0.6;2.2],m=2$,
\begin{gather*}
  \varphi(x)=\cos(18x-3)\sin(10x-7)+1.5, \\
  g_1(x)=\exp(-x/2)\sin(6x-1.5), \\
  g_2(x)=|x|\sin(2\pi x-0.5).
\end{gather*}
he exact solution of the problem is $x^*=2.0795,\varphi(x^*)=0.565$. The graphs of the functions $g_1(x),g_2(x)$,and $\varphi(x)$ are presented in Fig. \ref{fig:4_1}.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_1.png}
  \caption{One-dimensional global optimization problem}
  \label{fig:4_1}
\end{figure}
}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_2.png}
  \caption{A problem with partially computable functions}
  \label{fig:4_2}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_3.png}
  \caption{An ''index'' function}
  \label{fig:4_3}
\end{figure}

Since in the general case the problem \eqref{eq4:problem2} may have no solution (i.e., the feasible subdomain $Q_{m+1}$ might appear to be empty because of the incompatibility of the constrains) an auxiliary problem always having a solution is associated with it. Since conditions \eqref{eq4:condition} are equivalent to the conditions
\begin{displaymath}
  x\in Q_\nu,x\not\in Q_{\nu+1},
\end{displaymath}
this auxiliary problem can be written in the form
\begin{equation}
  \label{eq4:problem3}
  g_M^*=g_M(x_M^*)=\min\{g_M(x):x\in Q_M\},
\end{equation}
where $M$ is the greatest possible value of the index i. e.
\begin{equation}
  1\le M=\max\{\nu(x):x\in[a,b]\}\le m+1
\end{equation}
Since the set $Q_M$ is always nonempty, problem \eqref{eq4:problem3} always has a solution. When $M=m+1$ the solution $x^*=x^*_{m+1}$  is also the solution of the initial problem \eqref{eq4:problem}. When $M<m+1$ the inequality $g_M^*<0$ fulfilled necessarily can be used as an indicator of the incompatibility of the constraints.

@@

As a result, the arcs of $\psi(x)$ will be Lipschitzian with the constant $L=1$ in each subdomain $Q_\nu,1\le\nu\le M$. A graph of the function $\psi(x)$ for the example (\ref{ex4:problem}) is presented in Fig. \ref{fig:4_4}. This new function will have discontinuities of the first type at the boundary points of the subdomains $Q_\nu$ from \eqref{eq4:q}. Nevertheless, one can estimate the global minimizer $x^*_M$ using the results of $k$ trials for \eqref{eq4:trial} at the points $x_1,\dots, x_k$ in $[a,b]$ (see \cite{strongin1992}).

@@

For example, a case for $k=4$ is presented in Fig. \ref{fig:4_4}. The union of the segments highlighted with a bold line in Fig. \ref{fig:4_4} is a complement of the set from the right hand side of \eqref{eq4:lip_opt_est} and does not include the optimum point.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_4.png}
  \caption{An estimate of the optimum point}
  \label{fig:4_4}
\end{figure}

@@

This theorem is given without a proof, which can be found, for example, in \cite{stronginMarkin1986, stronginMarkin1987}. Various modifications of this algorithm and the corresponding theory of convergence are given in \cite{SergeyevFamularoPugliese2001,barkalovStrongin2002,SergeyevFamularoPugliese2003,sergeyev2006,sergeyevKhalafKvasov2007,sergeyevKvasovKhalaf2007}.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_5.png}
  \caption{Minimization by means of Penalty Function Method}
  \label{fig:4_5}
\end{figure}

@@

Here $k_i$ is the number of computations of the $i$ problem function, i.e., of the function $g_i(x )$.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_6.png}
  \caption{The minimization of an ''index'' function}
  \label{fig:4_6}
\end{figure}

\section{Index algorithm taking into account the existence of $\epsilon$-reserved solutions}
Let us return to the consideration of a one-dimensional problem of the kind \eqref{eq4:problem}
\begin{displaymath}
  \varphi(x^*)=\min\{\varphi(x):x\in [a,b], g_j(x)\le 0,1\le j\le m\}
\end{displaymath}
In the case when the objective function and the left-side parts of the constraints are partially computable, it is necessary to use the form \eqref{eq4:problem2}.
\example
{
Let us consider problem \eqref{eq4:problem} with $x\in[0,6,2,2]$, $m=3$ as an illustration:
\begin{gather*}
\varphi(x)=\cos(18x-3)\sin(10x-7)+1, \\
g_1(x)=\exp(-x/2)\sin(6x-1.5) \\
g_2(x)=\sin(4x-2.2)+\cos(6x-2.9) \\
g_3(x)=|x|\sin(2\pi x - 0.5).
\end{gather*}
}
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_7.jpg}
  \caption{The solution of a problem by Index Method}
  \label{fig:4_7}
\end{figure}
The functions $g_j , 1\le j\le 4$, are pictured in Fig. \ref{fig:4_7}. Assuming a partial calculability, the arcs of
the functions corresponding to the domains $Q_j ,1\le j\le 4$, from \eqref{eq4:q} are presented in Fig. \ref{fig:4_8}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_8.jpg}
  \caption{The problem solving by Index Method with the $\epsilon$-reservs}
  \label{fig:4_8}
\end{figure}

@@

As it has been demonstrated by the experiment with solving the example described above, the increase of the reserve values may result in an essential decrease of the trial concentration in the boundary points of non-feasible domains. The theoretical properties of the trial sequences generated by the index algorithm are determined by the following statement (see the proof in \cite{stronginSergeyev2000} also).
\begin{theorem}
  Assume the convergence conditions for the index algorithm to be satisfied and the inequalities
  \begin{displaymath}
    g_\nu(x)\ge\delta_\nu,x\in[\alpha,\beta],\delta_\nu\ge 0
  \end{displaymath}
  to be satisfied in a subinterval $[\alpha,\beta]\subset Q_\nu \: 1\le\nu\le m$, the length of which meets the condition
  \begin{displaymath}
    \beta-\alpha>\frac{2(r_\nu-1)(\delta_\nu+\epsilon_\nu)}{r^2_\nu L_\nu}.
  \end{displaymath}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_9.png}
  \caption{The operation characteristics in class $G_{SH}$}
  \label{fig:4_9}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_10.png}
  \caption{The operation characteristics in class $G_{HL}$}
  \label{fig:4_10}
\end{figure}

The operational characteristics for both method obtained for the test problem sets generated by $G_{SH}$ and $G_{HL}$ are presented in Figures \ref{fig:4_9} and \ref{fig:4_10}, respectively. The lower (solid) curves refer to the initial index algorithm while the upper (dashed) ones correspond to the index algorithm with the adaptive reserves. Such arrangement of the curves demonstrates the algorithm with the
adaptive reserves to provide much faster obtaining the estimates falling into given vicinity of the solution than the initial index algorithm. The average numbers of iterations for the method
without the reserves and for the method with the adaptive reserves were equal to 171 and 117 for the problems $G_{SH}$ and 199 and 142 for the ones $G_{HL}$ , respectively.

@@

\textbf{Proof} of the theorem is given in \cite{SergeyevFamularoPugliese2001}.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_11.jpg}
  \caption{The distribution of problems of the class $G_{SH}$}
  \label{fig:4_11}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_12.jpg}
  \caption{The distribution of problems of the class $G_{HL}$}
  \label{fig:4_12}
\end{figure}

The distribution of the interval lengths $[\alpha,\beta]$ from (4.41) including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig. \ref{fig:4_11}. Fig. \ref{fig:4_12} demonstrates a similar distribution for the problems of the second set. In these figures, the values $(\beta=\alpha)/(b-a)$ (in percentage) are plotted against the abscissa axes while the numbers of problems in the corresponding sets are plotted against the ordinate ones. In order to obtain the discussed estimates, the lengths of the intervals $[\alpha,\beta]$ corresponding to the respective functions were determined by a numerical method.

@@

The constraint functions $g_\nu( x ),\: 1 \le\nu\le 5$, and the objective function $\varphi (x)$ for a problem from the class $G_{HP}$ are plotted in Fig. \ref{fig:4_13} as an example. The feasible domains which are formed by the separate constraints are marked on the abscissa axis. The whole feasible domain of the problem (the intersection of all feasible domains of the constraints) is shown along with the graph of the objective function.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_13.png}
  \caption{Function plots of a test problem from the class $G_{HL}$}
  \label{fig:4_13}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_13.png}
  \caption{Operational characteristics for the class $G_{SH}$}
  \label{fig:4_14}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_15.png}
  \caption{Operational characteristics for the class $G_{HL}$}
  \label{fig:4_15}
\end{figure}

The operational characteristics for both the methods obtained on the sets produced by generators $G_{SH}$ and $G_{HL}$ are presented in figures \ref{fig:4_14} and \ref{fig:4_15}, respectively. The lower (solid) curves correspond to FCOM while the upper (dashed) ones --- to ACOM. These positions of the curves demonstrate that the algorithm with the adaptive check order provides in average faster obtaining the estimates falling into given vicinity of the solution than the algorithm with the fixed checking order.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_16.png}
  \caption{Average number of function values computations for the class $G_{SH}$}
  \label{fig:4_16}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/4_17.png}
  \caption{Average number of function values computations for the class $G_{HL}$}
  \label{fig:4_17}
\end{figure}

@@

  It follows from the trial execution rules and condition (4.35) that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive. Therefore, according to \eqref{eq4:z_const_par},
  \begin{displaymath}
    z_i=g_\nu(x_i)\ge z^*_nu,\:1\le\nu=\nu(x_i)\le m+1,\: 0\le i\le k.
  \end{displaymath}
    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit (4.56), taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality (4.57), one can derive out the truth of statement (4.58).
\end{proof}
\begin{lemma}
  \label{lm4:2}
  (on the existence of a feasible point). Assume the conditions of Theorem \ref{th4:convergence_par} to be fulfilled. Then, there exists the trial index $h\ge 1$, for which
  \begin{equation}
    \label{eq4:62}
    \nu(x^h)=m+1.
  \end{equation}
\end{lemma}
\begin{proof}
  Let us assume the opposite and consider the interval $[x_{t-1},x_t]$ including the problem solution $x^*$ from \eqref{eq4:problem} at the step $k$, i.e., $t=t(k)$. Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from (4.52) should satisfy the conditions
  \begin{equation}
    \label{eq4:63}
    x^*\in [\alpha,\beta]\subset[x_{t-1}, x_t].
  \end{equation}
  Then, from this condition as well as from the Lipschitz properties of the problem functions the relations
  \begin{gather}
    z_{t-1}\le L_\nu(\alpha - x_{t-1})\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t-1}) \\
    z_t\le L_\nu(x_{t}-\beta)\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t})
  \end{gather}
  follow. It was taken into account when deriving out of the above relations that
  \begin{displaymath}
    g_\nu(x)\le 0,\: x\in[\alpha,\beta]\subset Q.
  \end{displaymath}
  Analogously, the case $\nu=\nu(x_{t-1})=\nu(x_t)$ corresponds to the inequality
  \begin{equation}
    z_t+z_{t-1}\le L_\nu \Delta_t - L_\nu(\beta-\alpha).
  \end{equation}

  Now, from the characteristic rules \eqref{eq4:characteristic_par}, from inequalities (4.64)–(4.66), and from the inequality
  \begin{displaymath}
    z^*_\nu\ge-\epsilon_\nu
  \end{displaymath}
  resulting from \eqref{eq4:z_const_par} either the estimate
  \begin{displaymath}
    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+2(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
  \end{displaymath}
  or the estimate
  \begin{displaymath}
    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+4(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
  \end{displaymath}
  is true. Taking into account the third condition of Theorem \ref{th4:convergence_par} and the fourth one, in both cases we get the inequality
  \begin{equation}
    \label{eq4:67}
    R(t(k))>0
  \end{equation}
  which holds for sufficiently large step numbers $k$.

  From (4.58), \eqref{eq4:67}, and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from (4.63) follows. Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals. After that, at least one of those will include the point $x^*$. The repetition of this division generates (with $k\to\infty$) an infinite sequence of the subintervals shrinking to the point $x^*$ . In other words, if the third condition of Theorem \ref{th4:convergence_par} and the fourth one are satisfied, $x^*$ is the limit point of the sequence ${x^k}$.

@@

\begin{lemma}
  \label{lm4:3}
  on the feasibility of any limit point). Assume the conditions of Theorem \ref{th4:convergence_par} to be met. Then any limit point $\overline x$ of the sequence $\{x_k\}$ is feasible, i.e.,
  \begin{equation}
    \label{eq4:68}
    \nu(\overline x)=m+1
  \end{equation}
\end{lemma}

@@

  According to the trial execution rules and to conditions (4.35) and (4.72), for any interval from (4.70) the relations
  \begin{displaymath}
    \omega_1=\nu(x_{t-1})\in M,\: \omega_2=\nu(x_t)\in N,\:t=t(q^s),
  \end{displaymath}
  hold where $1\le \omega_1,\: \omega_2\le m$. Moreover, according to the rules (4.39) and (4.40) defining the constraints’ check order at trial points,
  \begin{displaymath}
    \nu(x_{t-1}),\:\nu(x_t)\in\{\omega_1,\omega_2\},\: t=t(q^{p+1}).
  \end{displaymath}
  In other words, for sufficiently large number $s>1$ the indices assigned to the boundary points of the considered system of intervals $[x_{t-1},x_t],\: t=t(q^s ): s\ge S$ can take the values from the set $\{\omega_1,\omega_2\}$ only.

@@

  I. Let us prove the first assertion of Theorem \ref{th4:convergence_par} based on Lemma \ref{lm4:2} and Lemma \ref{lm4:3}, i.e., that the point $x^*$ from \eqref{eq4:problem} is the limit point. Let us assume the opposite and consider two  possible cases. Assume $[x_{t-1},x_t],\: t=t(k)$ to be an interval including the point $x^*$ at the step $k$. Consider the condition
  \begin{equation}
    \label{eq4:75}
    \max\{\nu(x_{t-1}),\nu(x_t)\}=m+1
  \end{equation}
  for the indices of the boundary points of the given interval and assume that this condition is not satisfied for sufficiently large $k$. In this case, inclusions \eqref{eq4:63} should take place. This situation has been already considered in Lemma \ref{lm4:2}, where its incompatibility with the assumption of the point $x^*$ is not the limit point has been demonstrated.
